Title

### 2023-05-21 Cosas-Escritas-Antes.tex
Cambios

\newpage
\subsection{Unitary evolution $U = e^{iHt}$}

>>>>

\newpage
\subsection{Unitary evolution $U = e^{iHt}$}
\section{Exponential maps $e^{iHt}$} % {{{
\label{sec: Exponential maps}
\textbf{Puntos:}
\begin{itemize}
\item We will apply the results about 1PR circuit to another kind of operation, those given by $U  = e^{iHt}$ where $H$ is hermitian  and time independent.
\item \textbf{Lemma:}\textit{ The $n$-qubit operator $U^{iHt}$ can be implemented with a 1PR circuit of $n$ qubits if and only if $D = \text{diag}(e^{i\lambda_0 t}, \cdots, e^{i \lambda_{2^n-1}t})$ can be implemented with a 1PR circuit of $n$ qubits, where $\lambda_j$ are the eigenvalues of $H$.}
\item \textbf{Proof:}  Es b\'asicamente usar que $H = QDQ^{-1}$ para una matriz $Q$ que no depende de $t$ y por lo tanto no agrega compuertas parametrizadas al circuito. 
\item \textbf{Theorem:} \textit{ A n-qubit operator of the form $U = e^{iHt}$ can be implemented using a 1PR circuit if and only if the eigenvalues of $H$ are $-\lambda,0,\lambda$ (for $\lambda$ a real number) and the degeneracies of $\lambda$ and $-\lambda$ are both equal to $2^j$ for some $j\in \{0,1,\cdots,n-1\}$.}
\item \textbf{Proof:} Estoy buscando formas de simplificar la prueba, porque es medio larguita. Pero básicamente es usar el lema y entonces preocuparnos sólo por $D$. Luego  usar el teorema de la sección 3 y notar que sólo hay tres tipos de  $\lambda$s posibles y luego notar que deben de tener esas degeneraciones.
\item \textbf{Proof del regreso:} Aquí quiero recalcar que la prueba es constructiva, ya que consiste en mostrar cómo construir el circuito 1PR que hace la operación a partir de una $H$ que cumpla la hipótesis. 
\item \textbf{Examples.} Mencionar que por el teorema, los ejemplos son sistemas con 3 energías separadas por la misma distancia y con las degeneraciones dadas en el teorema. 
La verdad que no se me ocurren sistemas físicos comunes que cumplan eso.
\begin{itemize}
\item Cualquier matriz $H$ de un qubit cumple el teorema.
\item Cualquier matriz de $n$ qubits que sea el producto tensorial de matrices de Pauli para cada partícula (o sea algo como $\sigma_{i_1} \otimes \sigma_{i_2} \otimes \cdots \otimes \sigma_{i_n}$) cumple el teorema.
\item Estoy viendo para cadenas de espines si hay condiciones para las que cumplan las hipótesis del teorema, pero no he encontrado aún.
\end{itemize}
\item Hice explícitamente el circuito de 2 qubits para $H = \sigma_3 \otimes \sigma_2$ y lo simulé en las computadoras de IBM, obteniendo una fidelidad de $0.85$.
\end{itemize}
% }}}
\newpage

### 2023-05-21 main.tex
Cambios

% Use "Eq" instead of "Equation" for equation citations.
% }}}
\section{Introduction} % {{{

% }}}
\section{Quantum Channels}  \label{sec: Pauli Channels} % {{{
\cpnote{Es raro qeu la seccion y una subseccion se llamen igual. Por discutir con Tomas}
\tbnote{Lo cambié}
\cpnote{El texto esta bien, pero creo qeu la distribucino de subsecciones esta chistosna y 
los titulo tambien. Platicar con Tomas a ver a que llegamos. }

@@

\begin{figure}
\centering
\begin{subfigure}{0.9\textwidth}
\begin{subfigure}{.666\textwidth}
\centering
\includegraphics[width=.95\columnwidth]{images/tetra-def.png}
\end{subfigure}%
\begin{subfigure}{.333\textwidth}
\includegraphics[width=.95\columnwidth]{images/corte1-2.png} \\
\includegraphics[width=.95\columnwidth]{images/corte2-2.png} \\
\end{subfigure}
\includegraphics[width=.322\columnwidth]{images/corte3-2.png}
\includegraphics[width=.322\columnwidth]{images/corte4-2.png}
\includegraphics[width=.322\columnwidth]{images/corte5-2.png}
\end{subfigure}%
\begin{subfigure}{0.1\textwidth}
\includegraphics[width=.8\columnwidth]{images/col-2.png} \\ [2ex]
\end{subfigure}
\caption{\tbnote{Aquí está una forma alternativa en que puede quedar la figura. 
Usé $2(1-p)$ en lugar de $p$ para que se cumpla lo de
que $0$ es lo peor y $1$ lo mejor.}}
\end{figure}

@@

In general, producing this curve of states will require multiple rotations
parametrized by $p$. For instance, in the case of a two-qubit curve of states,
we would typically need three rotations that depend on $p$, similar to those
illustrated for the ancilla qubits in \fref{fig: circuit-pauli-1}.
Moreover, creating an $N$-qubit curve of states
usually necessitates $2^N-1$ parametrized rotations.
\cpnote{no me salen esas cuentas.} \tbnote{Se me fue el 2.}

@@

{\color{orange} In general, producing this curve of states for $N$ qubits will require
$2^N-1$ rotations parametrized by $p$,
such as the three rotations used for the ancilla
qubits in \fref{fig: circuit-pauli-1}.
However, it would be preferable to achieve the same effect using only one parametrized rotation. 
This would allow us to treat said rotation 
as a knob that smoothly traverses the curve of states.
Consequently, we are faced with the question of which curves of states, 
such as the one described in \eref{eq: parametrized state}, 
can be produced using just a single parametrized rotation. 
To clarify this, we provide the following definition 
for a circuit with one parametrized rotation.}

@@

Based on this definition, we aim to determine which curves of 
states can be generated using 1PR \cpnote{Si te parece bien el cambio 
en la sigla, cambialos todos} 
\tbnote{Sí me parece, los cambié} circuits. 
To accomplish this, we begin by proving that all 1PR circuits have 
{\color{orange} the form depicted in \fref{fig:1PR-circuit},
where the parametrized rotation is around $\sigma_3$ and
is applied to the last qubit.}

@@

\begin{theorem}
A 1PR circuit always implements an operator whose matrix,
in the computational basis, has the form:
\begin{eqnarray}
U = (\vec{v}_0, \vec{v}_1, \vec{v}_2, \cdots, \vec{v}_{2^n-1}),
\end{eqnarray}
where the column vectors are given by $\vec{v}_j = e^{is} \vec{a}_j + e^{-is} \vec{b}_j + \vec{c}_j$,
with $\vec{a}_j ,\vec{b}_j, \vec{c}_j$ orthogonal and $|\vec{a}_j|^2 + |\vec{b}_j|^2 + |\vec{c}_j|^2 = 1$.
\end{theorem}
\textbf{Proof:} 
\cpnote{Discutir la prueba, se ve demasiado larga y aburrida}
\tbnote{La cambié un poco, pero aún así.}
Denoting by $U$ the matrix for the 1PR circuit,
we can conclude from theorem 1 that $U=ARB$,
where $A$ and $B$ are unitary matrices and $R$ is a $\sigma_3$ rotation of angle $2s$
applied to the last qubit and controlled by some of the other
ones. 
We will calculate the $j$-th column of this matrix,
which is given by $U|j\rangle$.

First, applying $B$ to $|j\rangle$ results in $B|j\rangle = B_{0,j} |0\rangle + B_{1,j} |1 \rangle + \cdots + B_{2^n-1,j}|2^n-1\rangle$,
which can be rewritten by separating the
first $N-1$ qubits from the last one as:
\begin{equation}
B|j\rangle = \sum_{k=0}^{2^{N-1}-1} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j} |k\rangle |1 \rangle \right).
\end{equation}
After the operator $B$, we need to apply $R$.
To simplify the analysis, we separate the states of
the first $N-1$ qubits into those that fulfill the control conditions of the rotation
(which we denote as the set $\mathcal{C}$)
and those that do not, and write it as
\begin{eqnarray}
B|j\rangle = \sum_{k \in \mathcal{C}} \left(B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j} |k\rangle |1 \rangle \right) + \sum_{k \not\in \mathcal{C}} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j}|k\rangle |1\rangle \right),
\end{eqnarray}
Then, the rotation $R$ will only affect the states on the first sum (since they fulfill the control conditions)
and not the others. Therefore,  
remembering that a $\sigma_3$ rotation acts by adding a phase $e^{-is}$ to $|0\rangle$
and a phase $e^{is}$ to $|1\rangle$, $RB|j\rangle$ is:
\begin{small}
\begin{eqnarray}
&e^{-is} \sum_{k \in \mathcal{C}} B_{2k,j} |k\rangle |0\rangle + e^{is} \sum_{k \in \mathcal{C}} B_{2k+1,j} |k\rangle |1 \rangle 
+ \sum_{k \not\in \mathcal{C}} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1} |k\rangle |j \rangle \right)\\
& \qquad = e^{-is} \vec{b'}_j + e^{is} \vec{a'}_j + \vec{c'}_j,
\end{eqnarray}
\end{small}
where $\vec{a'} = \sum_{k \in \mathcal{C}} B_{2k,j}|k\rangle|0\rangle$, $\vec{b'}_j = \sum_{k \in \mathcal{C}} B_{2k+1,j} |k\rangle |1 \rangle$, $\vec{c'}_j =\sum_{k \not\in \mathcal{C}} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1} |k\rangle |j \rangle \right)$.
 These vectors are clearly orthogonal because they are each linear combinations of different 
orthogonal states of the computational basis. 
Moreover, they satisfy $|\vec{a'}_j|^2 + |\vec{b'}_j|^2 + |\vec{c'}_j|^2 = 1$ because this quantity is the norm of the $j$th column of $B$, 
which is unitary.

@@

This theorem implies that starting from the state $|0\rangle$,
the only curves of states we could expect to create
with a 1PR circuit are those of the form
\begin{equation}
|\eta(p)\rangle = \vec{c} + \vec{a} e^{is} + \vec{b} e^{-is},
\label{eq:curve-states}
\end{equation}
with 
\begin{equation}
|\vec{a}|^2 + |\vec{b}|^2 + |\vec{c}|^2 = 1  \; \text{and} \; \vec{a}, \vec{b}, \vec{c} \; \text{orthogonal vectors},
\label{eq:conditions-vecs}
\end{equation}
and $s = s(p)$ a function of $p$.
Furthermore, we can exhibit a 1PR circuit to create any of these curves of states.
To do it, we can use the circuit of \fref{fig:1PR-circuit} and set the parametrized
rotation $R$ on the last qubit to be controlled by all the other qubits..
Moreover, we define the matrices $A$ and $B$ as:
\begin{align*}
B = (\vec{d}, *,*, \cdots,*), \; 
A = (* , \cdots, *, \vec{c}/|\vec{c}|, \vec{b}/|\vec{b}|, \vec{a}/|\vec{a}|),
\end{align*}
where the column vector $\vec{d}$ is defined as 
$(0,0, \cdots,0 ,|\vec{c}|, |\vec{b}|, |\vec{a}|)^T$
and the asterisks can be filled up in any
way such that these matrices are unitary.
Then, starting from the initial state $|0\rangle = (1,0,0,\cdots,0)$
and applying the circuit (whose matrix is $ARB$), some direct calculations lead us  to 
the resulting curve of states
of \eref{eq:curve-states}.

@@

%\textbf{Puntos:}
%\begin{itemize}
%\item Use the last theorem to conclude directly which maps can be created using an 1PR circuit.
%\item \textbf{Examples:} Show examples that satisfy these conditions: Bit-Flip, parabolic maps, etc. 
%and an example that doesn't satisfy the conditions (ejemplo está en la tesis).
%\end{itemize}
% }}}
\section{Exponential maps $e^{iHt}$} % {{{
\label{sec: Exponential maps}
\textbf{Puntos:}
\begin{itemize}
\item We will apply the results about 1PR circuit to another kind of operation, those given by $U  = e^{iHt}$ where $H$ is hermitian  and time independent.
\item \textbf{Lemma:}\textit{ The $n$-qubit operator $U^{iHt}$ can be implemented with a 1PR circuit of $n$ qubits if and only if $D = \text{diag}(e^{i\lambda_0 t}, \cdots, e^{i \lambda_{2^n-1}t})$ can be implemented with a 1PR circuit of $n$ qubits, where $\lambda_j$ are the eigenvalues of $H$.}
\item \textbf{Proof:}  Es b\'asicamente usar que $H = QDQ^{-1}$ para una matriz $Q$ que no depende de $t$ y por lo tanto no agrega compuertas parametrizadas al circuito. 
\item \textbf{Theorem:} \textit{ A n-qubit operator of the form $U = e^{iHt}$ can be implemented using a 1PR circuit if and only if the eigenvalues of $H$ are $-\lambda,0,\lambda$ (for $\lambda$ a real number) and the degeneracies of $\lambda$ and $-\lambda$ are both equal to $2^j$ for some $j\in \{0,1,\cdots,n-1\}$.}
\item \textbf{Proof:} Estoy buscando formas de simplificar la prueba, porque es medio larguita. Pero básicamente es usar el lema y entonces preocuparnos sólo por $D$. Luego  usar el teorema de la sección 3 y notar que sólo hay tres tipos de  $\lambda$s posibles y luego notar que deben de tener esas degeneraciones.
\item \textbf{Proof del regreso:} Aquí quiero recalcar que la prueba es constructiva, ya que consiste en mostrar cómo construir el circuito 1PR que hace la operación a partir de una $H$ que cumpla la hipótesis. 
\item \textbf{Examples.} Mencionar que por el teorema, los ejemplos son sistemas con 3 energías separadas por la misma distancia y con las degeneraciones dadas en el teorema. 
La verdad que no se me ocurren sistemas físicos comunes que cumplan eso.
\begin{itemize}
\item Cualquier matriz $H$ de un qubit cumple el teorema.
\item Cualquier matriz de $n$ qubits que sea el producto tensorial de matrices de Pauli para cada partícula (o sea algo como $\sigma_{i_1} \otimes \sigma_{i_2} \otimes \cdots \otimes \sigma_{i_n}$) cumple el teorema.
\item Estoy viendo para cadenas de espines si hay condiciones para las que cumplan las hipótesis del teorema, pero no he encontrado aún.
\end{itemize}
\item Hice explícitamente el circuito de 2 qubits para $H = \sigma_3 \otimes \sigma_2$ y lo simulé en las computadoras de IBM, obteniendo una fidelidad de $0.85$.
\end{itemize}
% }}}
\section{Conclusion} % {{{

>>>>

% Use "Eq" instead of "Equation" for equation citations.
% }}}
\section{Introduction} % {{{
\tbnote{Poner intro luego}

% }}}
\section{Pauli channels and dynamical maps}  \label{sec: Pauli Channels} % {{{
\cpnote{Es raro qeu la seccion y una subseccion se llamen igual. Por discutir con Tomas}
\tbnote{Lo cambié}
\cpnote{El texto esta bien, pero creo qeu la distribucino de subsecciones esta chistosna y 
los titulo tambien. Platicar con Tomas a ver a que llegamos. }

@@

\begin{figure}
\centering
\begin{subfigure}{0.9\textwidth}
\begin{subfigure}{.666\textwidth}
\centering
\includegraphics[width=.95\columnwidth]{images/tetra-def.png}
\end{subfigure}%
\begin{subfigure}{.333\textwidth}
\includegraphics[width=.95\columnwidth]{images/corte1-2.png} \\
\includegraphics[width=.95\columnwidth]{images/corte2-2.png} \\
\end{subfigure}
\includegraphics[width=.322\columnwidth]{images/corte3-2.png}
\includegraphics[width=.322\columnwidth]{images/corte4-2.png}
\includegraphics[width=.322\columnwidth]{images/corte5-2.png}
\end{subfigure}%
\begin{subfigure}{0.1\textwidth}
\includegraphics[width=.8\columnwidth]{images/col-2.png} \\ [2ex]
\end{subfigure}
\caption{\tbnote{Aquí está una forma alternativa en que puede quedar la figura. 
Usé $2(1-p_{succ})$ en lugar de $p_{succ}$ para que se cumpla lo de
que $0$ es lo peor y $1$ lo mejor.}}
\end{figure}

@@

In general, producing this curve of states will require multiple rotations
parametrized by $p$. For instance, in the case of a two-qubit curve of states,
we would typically need three rotations that depend on $p$, similar to those
illustrated for the ancilla qubits in \fref{fig: circuit-pauli-1}.
Moreover, creating an $N$-qubit curve of states
usually necessitates $2^N-1$ parametrized rotations.
\cpnote{no me salen esas cuentas.} \tbnote{Se me olvidaba el 2, ya está bien}

@@

{\color{orange} In general, producing this curve of states for $N$ qubits will require
$2^N-1$ rotations parametrized by $p$,
such as the three rotations used for the ancilla
qubits in \fref{fig: circuit-pauli-1}.
However, it would be preferable to achieve the same effect using only one parametrized rotation. 
This would allow us to interpret said rotation 
as a knob that smoothly traverses the curve of states.
Consequently, we are faced with the question of which curves of states, 
such as the one described in \eref{eq: parametrized state}, 
can be produced using just a single parametrized rotation. 
To clarify this, we provide the following definition 
for a circuit with one parametrized rotation.}

@@

Based on this definition, we aim to determine which curves of 
states can be generated using 1PR \cpnote{Si te parece bien el cambio 
en la sigla, cambialos todos} 
\tbnote{Sí me parece, los cambié} circuits. 
To accomplish this, we begin by proving that all 1PR circuits have 
the form depicted in \fref{fig:1PR-circuit},
where the parametrized rotation is around $\sigma_3$ and
is applied to the last qubit.

@@

\begin{theorem}
A 1PR circuit always implements an operator whose matrix,
in the computational basis, has the form:
\begin{eqnarray}
U = (\vec{v}_0, \vec{v}_1, \vec{v}_2, \cdots, \vec{v}_{2^n-1}),
\end{eqnarray}
where the column vectors are given by $\vec{v}_j = e^{is} \vec{a}_j + e^{-is} \vec{b}_j + \vec{c}_j$,
with $\vec{a}_j ,\vec{b}_j, \vec{c}_j$ orthogonal and $|\vec{a}_j|^2 + |\vec{b}_j|^2 + |\vec{c}_j|^2 = 1$.
\end{theorem}
\textbf{Proof:} 
\cpnote{Discutir la prueba, se ve demasiado larga y aburrida}
\tbnote{La cambié un poco, pero no sé qué más cambiarle.}
Denoting by $U$ the matrix for the 1PR circuit,
we can conclude from theorem 1 that $U=ARB$,
where $A$ and $B$ are unitary matrices and $R$ is a $\sigma_3$ rotation of angle $2s$
applied to the last qubit and controlled by some of the other
ones. 
We will calculate the $j$-th column of this matrix,
which is given by $U|j\rangle$
and confirm that it has the form $\vec{v}_j$ as defined above.

First, applying $B$ to $|j\rangle$ results in $B|j\rangle = B_{0,j} |0\rangle + B_{1,j} |1 \rangle + \cdots + B_{2^n-1,j}|2^n-1\rangle$,
which can be rewritten by separating the
first $N-1$ qubits from the last one as:
\begin{equation}
B|j\rangle = \sum_{k=0}^{2^{N-1}-1} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j} |k\rangle |1 \rangle \right).
\end{equation}
After the operator $B$, the circuit applies the  controlled rotation $R$.
To simplify the analysis, we separate the states of
the first $N-1$ qubits into those that fulfill the control conditions of the rotation
(which we denote as the set $\mathcal{C}$)
and those that do not, and write it as
\begin{eqnarray}
B|j\rangle = \sum_{k \in \mathcal{C}} \left(B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j} |k\rangle |1 \rangle \right) + \sum_{k \not\in \mathcal{C}} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1,j}|k\rangle |1\rangle \right),
\end{eqnarray}
Then, the rotation $R$ will only affect the states on the first sum (since they fulfill the control conditions)
and not the others. Therefore,  
remembering that a $\sigma_3$ rotation acts by adding a phase $e^{-is}$ to $|0\rangle$
and a phase $e^{is}$ to $|1\rangle$, $RB|j\rangle$ is equal to:
\begin{small}
\begin{eqnarray}
&e^{-is} \sum_{k \in \mathcal{C}} B_{2k,j} |k\rangle |0\rangle + e^{is} \sum_{k \in \mathcal{C}} B_{2k+1,j} |k\rangle |1 \rangle 
+ \sum_{k \not\in \mathcal{C}} \left( B_{2k,j} |k\rangle |0\rangle + B_{2k+1} |k\rangle |j \rangle \right)\\
& \qquad = e^{-is} \vec{b'}_j + e^{is} \vec{a'}_j + \vec{c'}_j,
\end{eqnarray}
\end{small}
where $\vec{a'} = \sum_{k \in \mathcal{C}} B_{2k,j}|k\rangle|0\rangle$, $\vec{b'}_j = \sum_{k \in \mathcal{C}} B_{2k+1,j} |k\rangle |1 \rangle$, $\vec{c'}_j =\sum_{k \not\in \mathcal{C}} \big( B_{2k,j} |k\rangle |0\rangle+$ $ B_{2k+1} |k\rangle |j \rangle \big)$.
 These vectors are clearly orthogonal because they are each linear combinations of different 
orthogonal states of the computational base. 
Moreover, they satisfy $|\vec{a'}_j|^2 + |\vec{b'}_j|^2 + |\vec{c'}_j|^2 = 1$ because this quantity is the squared norm of the $j$th column of $B$, 
which is unitary.

@@

This theorem implies that starting from the state $|0\rangle$ (or any 
other starting state),
the only curves of states we could expect to create
with a 1PR circuit are those of the form $U|0\rangle$, that is, curves such as:
\begin{equation}
|\eta(p)\rangle = \vec{c} + \vec{a} e^{is} + \vec{b} e^{-is},
\label{eq:curve-states}
\end{equation}
with 
\begin{equation}
|\vec{a}|^2 + |\vec{b}|^2 + |\vec{c}|^2 = 1  \; \text{and} \; \vec{a}, \vec{b}, \vec{c} \; \text{orthogonal vectors},
\label{eq:conditions-vecs}
\end{equation}
and $s = s(p)$ a function of $p$.
Furthermore, given a curve of states as the one in
\eref{eq:curve-states}, we can always construct a 1PR circuit to create it.
One way of doing it, is to use the circuit of \fref{fig:1PR-circuit}, setting the parametrized
rotation $R$ on the last qubit to be controlled by all the other qubits
and defining the matrices $A$ and $B$ as
\begin{align*}
B = (\vec{d}, *,*, \cdots,*), \; 
A = (* , \cdots, *, \vec{c}/|\vec{c}|, \vec{b}/|\vec{b}|, \vec{a}/|\vec{a}|).
\end{align*}
Here the column vector $\vec{d}$ is defined as 
$(0,0, \cdots,0 ,|\vec{c}|, |\vec{b}|, |\vec{a}|)^T$
and the asterisks can be filled up in any
way such that these matrices are unitary.
Then, starting from the initial state $|0\rangle = (1,0,0,\cdots,0)$
and applying the circuit  of \fref{fig:1PR-circuit} (whose matrix representation is $ARB$), some direct calculations lead us  to 
the resulting curve of states
of \eref{eq:curve-states}.

@@

%\textbf{Puntos:}
%\begin{itemize}
%\item Use the last theorem to conclude directly which maps can be created using an 1PR circuit.
%\item \textbf{Examples:} Show examples that satisfy these conditions: Bit-Flip, parabolic maps, etc. 
%and an example that doesn't satisfy the conditions (ejemplo está en la tesis).
%\end{itemize}
% }}}

\section{Conclusion} % {{{
